---
title: "Proxmap Drafts"
runningheader: "Mapping public health resources" # only for pdf output
subtitle: "Mapping public health resources" # only for html output
date: "`r Sys.Date()`"
output:
  tufte::tufte_handout: default
---

```{r setup, include=FALSE}
library(tufte)
library(tidyverse)
library(tigris)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
options(tigris_use_cache = TRUE)
```

Let's start by importing the sample data:

```{r data-import, echo=FALSE}
freeclinics <- read.csv("data/Google_AZ_Free_Clinics.csv",encoding="UTF-8")
publichealth <- read.csv("data/Google_AZ_Public_Health.csv",encoding="UTF-8")
triballibraries <- read.csv("data/Google_AZ_Tribal_Libraries.csv",encoding="UTF-8")
# Right away gotta do some cleaning
# freeclinics$ADDRESS[1] <- "304 W Cedar St"
# Let's look at it
head(freeclinics,n=3)
head(publichealth,n=3)
head(triballibraries,n=3)
```
The non-ASCII characters (I see one in ADDRESS) are going to be a problem so let's convert now.
```{r no-unicode, echo=FALSE}
freeclinics$ADDRESS <- iconv(freeclinics$ADDRESS,"UTF-8","ASCII",sub="")
knitr::kable(freeclinics[1:3,])
```
No need to have the data in separate tables so let's combine them. But first, let's create a new variable to track the type of site.
```{r categorizing, echo=FALSE}
freeclinics %>% mutate(type="Free Clinic") %>% rename("ZIP"="ZIP5") -> freeclinics
publichealth %>% mutate(type="Public Health Office") %>% rename("ZIP"="ZIP.CODE") -> publichealth
triballibraries %>% mutate(type="Tribal Library",STATE="AZ") %>%
  rename("NAME"="LIBNAME") %>% select(NAME,ADDRESS,CITY,STATE,ZIP,type) -> triballibraries
sites <- rbind(freeclinics,publichealth,triballibraries)
sites[1,]
```
And now let's check our data types and get rid of inconsistent capitalization:
```{r cleanup, echo=FALSE}
sites$type <- as.factor(sites$type)
sites$NAME <- str_to_title(sites$NAME)
sites$ADDRESS <- str_to_title(sites$ADDRESS)
sites$CITY <- str_to_title(sites$CITY)
as_tibble(sites)
```
Ok. Now that the data are neat, we can get on to the fun stuff...

There are a couple of simple ways to map the great state of Arizona:
```{r azmaps, echo=TRUE, fig.cap="Simple map of Arizona"}
usmap <- states() # Download USA maps deom US Census Bureau
azmap <- usmap[usmap$NAME=="Arizona",] # Limit ourselves to AZ
plot(azmap, max.plot=1)
```
```{r zipmap, echo=TRUE, fig.cap="Map of Arizona ZIP codes"}
azzipsmap <- zctas(year=2010,state="Arizona")
plot(azzipsmap, max.plot=1)
```
It gets even more interesting (Bless the US Census Bureau!)
```{r AZnativeareas, echo=TRUE, fig.cap="Native Areas in Arizona", message=FALSE, warning=FALSE}
nat <- native_areas()
AZnat <- sf::st_intersection(azmap,nat)
plot(AZnat, max.plot=1)
```
```{r AZtribalcensus, echo=TRUE, fig.cap="Tribal Census tracts in Arizona", message=FALSE, warning=FALSE}
tribcen <- tribal_census_tracts()
AZtrib <- sf::st_intersection(azmap,tribcen)
plot(AZtrib, max.plot=1)
```

To be continued!!!


